---
title: "One trillion trees analysis"
author: "Matt Russell"
date: "8/4/2020"
output: html_document
---

# Import and prep the data

Load the packages:

```{r}
library(googlesheets4)
library(tidyverse)
library(quantreg)
```

Read in the data:

```{r}
tree <- read_sheet("https://docs.google.com/spreadsheets/d/16kqmkcDnHI7OnyvSL72iX5BxdWAchsPZMS-CA5Ek3Yk/edit?usp=sharing")

# Remove AK and HI 
tree <- tree %>% 
  filter(!state %in% c("AK", "HI"))

state <- read_sheet("https://docs.google.com/spreadsheets/d/1fnquIcYugmQfIZNhe2Wgeih03jfXZZOR-A9Tw2i3EwE/edit?usp=sharing")
# Remove AK and HI 

state <- state %>% 
  filter(!state %in% c("AK", "HI"))
```

Calculate how many trees across US, in each diameter class, and in each US state:

```{r}
tree %>% 
  summarize(num_trees_total_us = sum(n_trees))

tree %>% 
  group_by(size) %>% 
  summarize(num_trees_total_us = sum(n_trees))

tree.summ <- tree %>% 
  group_by(state) %>% 
  summarize(num_trees_total = sum(n_trees))
tree.summ
```

Merge tree and state and calculate per-hectare values:

```{r}
state <- state %>% 
  mutate(prop_forest = forest_area_ac/total_area_ac)

tree <- tree %>% 
  inner_join(state) %>% 
  mutate(forest_area_ha = forest_area_ac * 0.404686,
         forest_area_ha_MM = (forest_area_ac * 0.404686) / 1000,
         num_trees_total_MMM = n_trees  / 1000000)

tree.summ <- tree.summ %>% 
  inner_join(state) %>% 
  mutate(forest_area_ha = forest_area_ac * 0.404686,
         forest_area_ha_MM = (forest_area_ac * 0.404686) / 1000,
         num_trees_total_MMM = num_trees_total / 1000000000)

```

Make graph of quantile regression lines:

```{r}
p.trees <- ggplot(tree.summ, aes(forest_area_ha_MM, num_trees_total_MMM)) +
  geom_point() +
#  stat_smooth(method = "lm", col = "red", se=F)+
  geom_quantile(quantiles = c(0.5, 0.79, 0.99)) +
  labs(x = "Forest land in US state (million hectares)",
       y = "Trees (billions)") +
  annotate("text", x = 13.5, y = 13, label = "Quantile = 0.50")+
  annotate("text", x = 13.5, y = 55, label = "Quantile = 0.99")+
  annotate("text", x = 13.5, y = 30, label = "Quantile = 0.79")+
  theme(panel.background = element_rect(fill = "NA"),
        axis.line = element_line(color = "black"),
        axis.text = element_text(color = "black"))

p.trees
```

```{r, eval = F}
ggsave(filename = "Fig1.pdf", plot = p.trees, width = 5, height = 3, 
       units = "in", scale = 1)
```

Model bumber of trees at different quantiles:

```{r}
# Simple linear regression

lm.all <- lm(num_trees_total_MMM ~ forest_area_ha_MM, data = tree.summ)
summary(lm.all)

# Quantile regression at 99th quantile
qr.all <- rq(num_trees_total_MMM ~ forest_area_ha_MM, data = tree.summ, tau = c(0.5, 0.99))
qr.all

tree.summ <- tree.summ %>% 
  mutate(num_trees_total_99 = -0.2315100 + (3.3475 * forest_area_ha_MM))

tree.summ %>% 
  summarize(num_trees_total_us = sum(num_trees_total_MMM),
            num_trees_total_99_us = sum(num_trees_total_99))

# Search for quantile that gives the US 76 billion more trees (431 total trees)
# Quantile regression at 79th quantile
qr.all <- rq(num_trees_total_MMM ~ forest_area_ha_MM, data = tree.summ, tau = c(0.79))
qr.all

tree.summ <- tree.summ %>% 
  mutate(num_trees_total_79 = -0.10225 + (1.72964 * forest_area_ha_MM))

tree.summ %>% 
  summarize(num_trees_total_us = sum(num_trees_total_MMM),
            num_trees_total_79_us = sum(num_trees_total_79))

# Quantile regression at 50th quantile
qr.all <- rq(num_trees_total_MMM ~ forest_area_ha_MM, data = tree.summ, tau = c(0.50))
qr.all

tree.summ <- tree.summ %>% 
  mutate(num_trees_total_50 = -0.10225 + (1.72964 * forest_area_ha_MM))

tree.summ %>% 
  summarize(num_trees_total_us = sum(num_trees_total_MMM),
            num_trees_total_50_us = sum(num_trees_total_50))

# By size class

# Quantile regression at 99th quantile - saplings
tree.sap <- tree %>% 
  filter(size == "Sapling")

qr.sap <- rq(num_trees_total_MMM ~ forest_area_ha_MM, data = tree.sap, tau = c(0.99))
qr.sap
summary(qr.sap)

tree.sap <- tree.sap %>% 
  mutate(num_sap_total_99 = -225.2588 + (2876.787 * forest_area_ha_MM))

tree.sap %>% 
  summarize(num_sap_total_us = sum(num_trees_total_MMM),
            num_sap_total_99_us = sum(num_sap_total_99))

# Quantile regression at 79th quantile - saplings
qr.sap <- rq(num_trees_total_MMM ~ forest_area_ha_MM, data = tree.sap, tau = c(0.79))
qr.sap
summary(qr.sap)

tree.sap <- tree.sap %>% 
  mutate(num_sap_total_79 = -101.5835 + (1306.3957 * forest_area_ha_MM))

tree.sap %>% 
  summarize(num_sap_total_us = sum(num_trees_total_MMM),
            num_sap_total_79_us = sum(num_sap_total_79))

# Quantile regression at 99th quantile - poletimber
tree.pole <- tree %>% 
  filter(size == "Poletimber")

qr.pole <- rq(num_trees_total_MMM ~ forest_area_ha_MM, data = tree.pole, tau = c(0.99))
qr.pole
summary(qr.pole)

tree.pole <- tree.pole %>% 
  mutate(num_pole_total_99 = -12.22 + (364.28 * forest_area_ha_MM))

tree.pole %>% 
  summarize(num_pole_total_us = sum(num_trees_total_MMM),
            num_pole_total_99_us = sum(num_pole_total_99))

# Quantile regression at 79th quantile - poletimber
qr.pole <- rq(num_trees_total_MMM ~ forest_area_ha_MM, data = tree.pole, tau = c(0.79))
qr.pole
summary(qr.pole)

tree.pole <- tree.pole %>% 
  mutate(num_pole_total_79 = -1.00285 + (276.37782 * forest_area_ha_MM))

tree.pole %>% 
  summarize(num_pole_total_us = sum(num_trees_total_MMM),
            num_pole_total_79_us = sum(num_pole_total_79))

# Quantile regression at 99th quantile - sawtimber
tree.saw <- tree %>% 
  filter(size == "Sawtimber")

qr.saw <- rq(num_trees_total_MMM ~ forest_area_ha_MM, data = tree.saw, tau = c(0.99))
qr.saw
summary(qr.saw)

tree.saw <- tree.saw %>% 
  mutate(num_saw_total_99 = -6.79 + (193.814 * forest_area_ha_MM))

tree.saw %>% 
  summarize(num_saw_total_us = sum(num_trees_total_MMM),
            num_saw_total_99_us = sum(num_saw_total_99))

# Quantile regression at 79th quantile - sawtimber
qr.saw <- rq(num_trees_total_MMM ~ forest_area_ha_MM, data = tree.saw, tau = c(0.79))
qr.saw
summary(qr.saw)

tree.saw <- tree.saw %>% 
  mutate(num_saw_total_79 = 6.46502 + (142.26666 * forest_area_ha_MM))

tree.saw %>% 
  summarize(num_saw_total_us = sum(num_trees_total_MMM),
            num_saw_total_79_us = sum(num_saw_total_79))
```



```{r}
dist <- tribble(
  ~`Size class`, ~`Time`, ~`Billion trees`,
  "Saplings", "Current day", 261,
  "Poletimber", "Current day", 61,
  "Sawtimber", "Current day", 32,
  "Saplings", "Quantile = 0.79", 331,
  "Poletimber", "Quantile = 0.79", 71,
  "Sawtimber", "Quantile = 0.79", 36,
  "Saplings", "Quantile = 0.99", 729,
  "Poletimber", "Quantile = 0.99", 93,
  "Sawtimber", "Quantile = 0.99", 49
)

p.trees.size2 <- ggplot(dist, aes(reorder(`Size class`, -`Billion trees`), `Billion trees`, fill = factor(`Time`))) +
  geom_bar(stat = "identity", position = "dodge2", col = "black") +
  labs(x = "Tree size",
       y = "Trees (billions)") +
  scale_fill_brewer(palette="Dark2") +
  theme(panel.background = element_rect(fill = "NA"),
        axis.line = element_line(color = "black"),
        axis.text = element_text(color = "black"),
        legend.position = c(0.75, 0.75),
        legend.title = element_blank())

p.trees.size2
```

```{r, eval = F}
ggsave("Fig2.pdf", plot = p.trees.size2, width = 5, height = 3, 
       units = "in", scale = 1)
```

Enter data on country land area. Data from [Food and Agriculture Organization of the United Nations](http://www.fao.org/faostat/en/#home) 

```{r}
# Enter data

country <- tribble(
~Country, ~Forestland,
"Russian Federation", 815,
"Brazil", 494,
"Canada", 347,
"United States of America", 310,
"China", 208,
"Democratic Republic of the Congo", 153,
"Australia", 125,
"Indonesia", 91,
"Peru", 74,
"India", 71,
"Mexico", 66,
"Colombia", 59,
"Angola", 58,
"Bolivia", 55,
"Zambia", 49,
"All other countries", 1080
)

country %>% 
  mutate(Total_area = sum(Forestland))

country2 <- country %>% 
 mutate(Forestland_pct=Forestland/sum(Forestland))

country2 
```



